Collocation and keyword analysis on India-related text of *Natural History*

### 2023-08-27 paper_quarto.qmd
0827 final refine

---
engine: knitr
title: "Collocation and keyword analysis on India-related text of *Natural History*"
author: "Dawn, Lizao Zhuang (r0914937)"
format: 
  html:
    theme: journal
    toc: true
    code-link: true
    abstract-title: "Abstract"
number-sections: true
echo: false
warning: false
tbl-cap-location: top
fig-cap-location: top
bibliography:
  - references.bib
  - packages.bib  
csl: unified-style-sheet-for-linguistics.csl
abstract: 'This paper presents a comprehensive analysis of textual collocations and keywords related to "India" in a corpus of *Natural History*, the earliest known encyclopedia written by Pliny the Elder in the 1st century AD. By analyzing these word associations and significant terms, this study seeks to identify the main themes and topics related to the Indian region within the text. This will shed light on how Pliny depicts and imagines this geographical area, providing insights into the spatial representation and imagination in the work.'
---

@@

refcorpus_folder <- here("NH_wholetext")
fnames_wholetext <- get_fnames(refcorpus_folder) %>% 
  keep_re("[.]txt")

@@

# Exclude short file names with the same name from target corpus in the reference corpus
fnames_wholetext_short_filtered <- fnames_wholetext_short %>%
  setdiff(fnames_indiatext_short)

# Filter full file names based on filtered short names
fnames_wholetext_filtered <- fnames_wholetext[fnames_wholetext_short %in% fnames_wholetext_short_filtered]

@@

During the extraction process of the corpus files, the texts are already converted to lower case.

@@

Textual collocation refers to the observation of some words frequently occurring with the target word in the context. Different from the surface collocation, which taking the collocating span, i.e. the certain words left and right to the target word into consideration [@evert2009], the word span in this case refers to the textual unit around the target word. Setting "India" as the target word in the corpus of all paragraphs in *Natural History*, the `mclm::text_cooc()` function returns the frequency of words appeared in the context of "India". As mentioned in the Methodology Section (@sec-method), the returned scores are filtered with a statistical criteria for gaining a list of the most frequently observed words in the context of word "India" with strong significance.

The prominent textual collocates of "India" identified with this approach suggests the potential themes about Indian region in the narrative of Pliny.

@@

```{r}
coocs <- fnames_wholetext %>% 
  text_cooc("(?xi)  ^ india $")

@@

theme_set(theme_minimal(base_size = 15))
g1 <- top_scores_colloc_df %>%
  ggplot(aes(x = G_signed, y = a)) +
  labs(x = "signed $G^2$", y = "Absolute frequency") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

@@

Additionally, in order to examine the significant phrases, i.e. "keyword", in the India-related content in the work, as a comparison and supplement to the conclusion of textual collocation analysis, a keyword analysis is employed for the India-related text.

The frequency lists for the **target corpus** (India-related texts) and **reference corpus** (whole text) are generated respectively, and the association scores are calculated through the `mclm::assoc_scores()` function. Similar to the collocation analysis, the returned scores are filtered with a criterion of PMI \>= 2 and signed $G^2$ \>= 3.84, to ensure the selected words for further discussion have a strong attraction to the target corpus with statistical significance [@montes2022].

@@

```{r}
#| label: keyword

@@

top_scores_kw %>% # also valid for top_scores_colloc
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>%         # sort by PMI (descending) 
  slice_max(order_by = G_signed, n = 20, with_ties = TRUE) %>%   
  kbl(col.names = c("Type", "Frequency", "PMI", r"(Signed $G^2$)")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")
```

@@

theme_set(theme_minimal(base_size = 15))
g <- top_scores_kw_df %>%
  ggplot(aes(x = G_signed, y = a)) +
  labs(x = "signed $G^2$", y = "Absolute frequency") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

@@

**Geographical connections**: india, arabia, indian, ganges, indus, ethiopia, taprobane(G), thence, sail(A)

**Animal/Plant**: elephants, pepper(A)

**Precious stone**: stones, amber, rock-crystal, glass, beryls(G)

**Historical figure**: alexander, ichthyophagi(G), megasthenes(G)

**Ethnic group**: indians, tribe(A)

**Descriptive expression (numeric)**: hundred, thousand, fifty(A)

@@

```{r}
#| label: tbl-kw_frequency_compare
#| tbl-cap:  Comparison of overlapping keywords in India-related text sorted by absolute frequency and G signed with that of reference corpus
flist_ref_df <- as_tibble(flist_ref)
flist_target_df <- as_tibble(flist_target)

@@

tar_nrm_freq_lookup <- setNames(flist_target_df$nrm_freq, flist_target_df$type)

nrm_freq_lookup <- setNames(flist_ref_df$nrm_freq, flist_ref_df$type)

results <- tibble(
  Word = filtered_top_scores$type,
  Target_Nrm_Frequency = tar_nrm_freq_lookup[filtered_top_scores$type],
  Reference_Nrm_Frequency = nrm_freq_lookup[filtered_top_scores$type]
) %>% 
  arrange(desc(Target_Nrm_Frequency)) %>%
  kbl(col.names = c("Type", "Target_Nrm_Frequency", "Reference_Nrm_Frequency")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")

results

@@

```{r}
#| label: tbl-kw_frequency_pmi
#| tbl-cap:  Comparison of overlapping keywords in India-related text sorted by absolute frequency and G signed with that of reference corpus

@@

results1 <- tibble(
  Word = filtered_top_scores_pmi$type,
  Target_Nrm_Frequency = tar_nrm_freq_lookup[filtered_top_scores_pmi$type],
  Reference_Nrm_Frequency = nrm_freq_lookup[filtered_top_scores_pmi$type]
)%>% 
  arrange(desc(Target_Nrm_Frequency)) %>%
  kbl(col.names = c("Type", "Target_Nrm_Frequency", "Reference_Nrm_Frequency")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")

results1

@@

keyword_PMI_list %>% 
  as_tibble() %>%
  select(Word, Exp, Category, Remark) %>%                        
  kbl(col.names = c("Word", "Exp", "Category", "Remark")) %>% 
  kable_minimal()
```

@@

# Plot the tag distribution
ggplot(tag_counts, aes(x = n, y = reorder(Category, -n), fill = Category)) +
  geom_bar(stat = "identity") +
  labs(x = "Count", y = "Category", fill = "Category") +
  theme_minimal() +
  coord_flip()
```

@@

Furthermore, the examination of keywords in India-related text aligns with the observations of the collocation analysis and highlights potential topics on stones and rivers about India in the work. These findings enhance the interpretation of the identified themes, such as geographical connectivity and merchandise trade, and also reflect Pliny's general geographical concern in the historical context of ancient Rome.

As this paper serves as a preliminary exploration, it encourages further in-depth investigations of the text. A comparative analysis of collocation and keywords centering other neighboring regions, and a closer examination of "stone" or "river" in the context, incorporating additional literature regarding the historical and cultural background may offer greater insights and enrich the discussion.

>>>>

---
engine: knitr
title: "Collocation and keyword analysis on India-related text of *Natural History*"
author: "Dawn, Lizao Zhuang (r0914937)"
format: 
  html:
    theme: journal
    toc: true
    code-link: true
    abstract-title: "Abstract"
number-sections: true
echo: false
warning: false
tbl-cap-location: top
fig-cap-location: top
bibliography:
  - references.bib
  - packages.bib  
csl: unified-style-sheet-for-linguistics.csl
abstract: 'This paper presents a comprehensive analysis of textual collocations and keywords related to "India" in a corpus of *Natural History*, the earliest known encyclopedia written by Pliny the Elder in the 1st century AD. By analyzing these word associations and significant terms, this study seeks to identify the main themes and topics related to the Indian region within the text. The investigation of how Pliny depicts India in *Natural Hisotry* could potentially contribute to the discussion of the spatial representation and imagination in the work.'
---

<script type="text/x-mathjax-config">MathJax.Hub.Config({tex2jax: {inlineMath: [["$","$"]]}})</script>
<script async src="https://mathjax.rstudio.com/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML"></script>

@@

# read file names of the corpra
refcorpus_folder <- here("NH_wholetext")
fnames_wholetext <- get_fnames(refcorpus_folder) %>% 
  keep_re("[.]txt")

@@

# exclude short file names with the same name from target corpus in the reference corpus
fnames_wholetext_short_filtered <- fnames_wholetext_short %>%
  setdiff(fnames_indiatext_short)

# filter full file names based on filtered short names
fnames_wholetext_filtered <- fnames_wholetext[fnames_wholetext_short %in% fnames_wholetext_short_filtered]

@@

During the extraction process of the corpus files, the texts are already converted to lower case. In addition, the closing quotation mark "'", which ranks high while provides limited information in the initial observation of the returned frequency lists, is replaced by an empty string for a more informative return for further discussions.

@@

@@

```{r}
#| lable: obtain and filter collocation association scores

coocs <- fnames_wholetext %>% 
  text_cooc("(?xi)  ^ india $")

@@

theme_set(theme_minimal(base_size = 15))
g1 <- top_scores_colloc_df %>%
  ggplot(aes(x = G_signed, y = a)) +
  labs(x = r"(Signed G)", y = "Absolute frequency") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

@@

@@

```{r}
#| label: obtain and filter keywords association scores

@@

top_scores_kw %>% 
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>%         # sort by PMI (descending) 
  slice_max(order_by = G_signed, n = 20, with_ties = TRUE) %>%   
  kbl(col.names = c("Type", "Frequency", "PMI", r"(Signed $G^2$)")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")
```

@@

theme_set(theme_minimal(base_size = 15))
g <- top_scores_kw_df %>%
  ggplot(aes(x = G_signed, y = a)) +
  labs(x = r"(Signed G)", y = "Absolute frequency") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

@@

**Geographical connections**: *india*, *arabia*, *indian*, *ganges*, *indus*, *ethiopia*, *taprobane*(G), *thence*, *sail*(A)

**Animal/Plant**: *elephants*, *pepper*(A)

**Precious stone**: *stones*, *amber*, *rock-crystal*, *glass*, *beryls*(G)

**Historical figure**: *alexander*, *ichthyophagi*(G), *megasthenes*(G)

**Ethnic group**: *indians*, *tribe*(A)

**Descriptive expression (numeric)**: *hundred*, *thousand*, *fifty*(A)

@@

```{r}
#| label: tbl-kw_frequency_compare
#| tbl-cap:  Normalized frequency comparison of overlapping top 20 ranking keywords in India-related text sorted by absolute frequency and signed $G^2$ score in target and reference corpus
flist_ref_df <- as_tibble(flist_ref)
flist_target_df <- as_tibble(flist_target)

@@

tar_nrm_freq_lookup <- setNames(flist_target_df$nrm_freq, flist_target_df$type)
nrm_freq_lookup <- setNames(flist_ref_df$nrm_freq, flist_ref_df$type)

results_keyword <- tibble(
  Word = filtered_top_scores$type,
  Target_Nrm_Frequency = tar_nrm_freq_lookup[filtered_top_scores$type],
  Reference_Nrm_Frequency = nrm_freq_lookup[filtered_top_scores$type]
) %>% 
  arrange(desc(Target_Nrm_Frequency)) %>%
  kbl(col.names = c("Type", "Target_Nrm_Frequency", "Reference_Nrm_Frequency")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")

results_keyword

@@

```{r}
#| label: tbl-kw_frequency_pmi
#| tbl-cap:  Normalized frequency comparison of top 20 ranking keywords in India-related text sorted by PMI in target and reference corpus

@@

results_keyword_pmi <- tibble(
  Word = filtered_top_scores_pmi$type,
  Target_Nrm_Frequency = tar_nrm_freq_lookup[filtered_top_scores_pmi$type],
  Reference_Nrm_Frequency = nrm_freq_lookup[filtered_top_scores_pmi$type]
)%>% 
  arrange(desc(Target_Nrm_Frequency)) %>%
  kbl(col.names = c("Type", "Target_Nrm_Frequency", "Reference_Nrm_Frequency")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")

results_keyword_pmi

@@

keyword_PMI_list %>% 
  as_tibble() %>%
  select(Word, Exp, Category, Remark) %>%                        
  kbl(col.names = c("Word", "Explaination", "Category", "Remark")) %>% 
  kable_minimal()
```

@@

# plot the tag distribution
ggplot(tag_counts, aes(x = n, y = reorder(Category, -n), fill = Category)) +
  geom_bar(stat = "identity") +
  labs(x = "Count", y = "Category", fill = "Category") +
  theme_minimal() +
  coord_flip()
```

@@

Moreover, the examination of keywords in India-related text aligns with the observations of the collocation analysis and highlights potential topics on stones and rivers about India in the work. These findings enhance the interpretation of the identified themes, such as geographical connectivity and merchandise trade, and also reflect Pliny's general geographical concern in the historical context of ancient Rome.

As this paper serves as a preliminary exploration, it encourages further in-depth investigations of the text. A comparative analysis of collocation and keywords centering other neighboring regions to India, and a closer examination of "stone" or "river" in the context, incorporating additional literature regarding the historical and cultural background may offer greater insights and enrich the discussion.
